---
title: 'Exercise 1: Supermarket data'
author: "??YOUR NAME??"
date: "June 22, 2017"
output: 
  html_document:
    theme: sandstone
    code_folding: show
---

The *Data Wrangling Cheat Sheet* and the *Data Visualization Cheat Sheet* may be helpful. 

This exercise is divided into four parts. The first part is on importing data and getting 
acquainted with the use of an R Markdown document and R chunks for writing R code. The 
second part is on using ggplot2 for graphics. The third part is on using some more advanced
features from the dplyr package for working with data in R. The last part is 
on the customization of the report and the output of chunks. 

This document is available as the file `Exercise_1_Data.Rmd` as well as the `Exercise_1_Data.html`. 
You should start by opening the Rmd-file in RStudio. If you then click the *Knit* button, you will 
get the HTML result shown in the Viewer pane in RStudio. 

```{r init}
library(tidyverse)
theme_set(theme_bw())
```

## Exercise 1A. Working with an R Markdown document {#1A}

### Reading data

The data for this exercise is available from 

[http://math.ku.dk/~richard/Annalect/Supermarket.txt](http://math.ku.dk/~richard/Annalect/Supermarket.txt)

* Start by importing this data set using RStudio. You can either do this via the *Import Dataset* button
that you find in the Environment pane, or from the file menu. Either way, this file is a CSV file. 
* Before importing the data, change the columns containing the ItemId and CatId from integer to character,
and the Week number from character to integer. Note the Code Preview. Highlight and copy the code before clicking on *Import*.

Once you have imported your data set, it appears in your Global Environment (the Environment pane) 
as *Supermarket*. If you click on the small table icon, you get a preview of the data. 

* Now paste the code that you copied before into the empty code chunk below for reading in the supermarket data. Before you do anything else, delete the line `View(Supermarket)` from the pasted code.

```{r reading_data, cache=TRUE}
## This is an empty code chunk. 
```

When you read in a data set using the `read_csv` function from the `readr` package, you will get a 
data table called a *tibble* in R. The chunk above has a cache option so that the data set is not read in
repeatedly from the url. 

The data set contains sales for a major supermarket chain in Sweden from the first 12 weeks of 2007. Most 
variable names are self-explanatory. The four marketing variables are indicators of whether the different marketing 
strategies have been used. The variable `NormalSale` is the expected (seasonally adjusted) number of sold items.
The variable `TotalSale` is the actual number of items sold. 

### Cleaning, filtering and transforming data

This first code chunk counts the number of observations with `DiscountSEK` missing. Once you have read in 
the data, you can evaluate it in the R console by pressing the small "play" button to the right of the chunk. 
Start by evaluating the very first code chunk in this document called `init`. It will load the tidyverse 
packages. As you go through the exercise, remove the chunk option `eval=FALSE` from the chunks in this document.

```{r na_discount, eval=FALSE}
filter(Supermarket, is.na(DiscountSEK)) %>% 
  count()
```

* What is the keyboard shortcut for evaluating a chunk?
* Change the code so that it counts the number of observations with `DiscountSEK` either negative or missing.
* Insert a new chunk where you discard all observations with a price less than 1 SEK and replace negative and missing values of DiscountSEK by 0 for the remaining observations. 
* Construct a new variable `NormalPrice` containing the price without discount.
* If time permits, count how many observations that fall in each category and construct a cross-tabulation between `Category` and `Week`.

## Exercise 1B. Basic plotting of data using geoms and aesthetics {#1B}

The following code chunk subsets the data to five specific items. They all belong to the 
category *Fisk & Skaldjur*.

```{r fisk, eval=FALSE, cache=TRUE}
fishItems <- c("7311170031149", "7311171002704", "7311171003510", "7340011391968", "7393835042264")
Fish <- filter(Supermarket, ItemId %in% fishItems)
```

* Using the `Fish` data, construct a barplot that shows the number of items sold each week.
* Add a fill aesthetic to distinguish the different items
* Plot the total sale against the normal sale.  
* Add a color aesthetic to distinguish the different items. 
* Add a line with slope 1 and intercept 0 using `geom_abline`. Give an interpretation.
* Transform both axes using a square root transformation.
* The relative total sale is `TotalSale / NormalSale`. Plot the relative total sale against the week number,
still using colors to distinguish the different items.

For the last question above, you may want to experiment with the following:

* Use of a logarithmic second axis.
* Use of `position = position_jitterdodge(0)` as an argument to `geom_point`.
* Replace `geom_point` by `geom_boxplot` and the `color` aesthetic with the `fill` aesthetic. For
this to work, it is also necessary to convert `Week` into a factor by `factor(Week)`.

## Exercise 1C: Using dplyr and ggplot in combination {#1C}

We first construct two data tables from the full data set. 

```{r fishSize, eval = FALSE, cache=TRUE}
storeSize <- Supermarket %>% group_by(StoreId) %>% 
  summarize(size = median(NormalSale))
Fish <- filter(Supermarket, CatId == "423230")
```

The `storeSize` table contains the stores and a proxy variable constructed 
so as to quantify the size of the store in terms of median sold items. The 
`Fish` table contains all observations belonging to the category *Fisk & Skaldjur*.

* Construct a table from `Fish` that contains the number of distinct items sold by each store. 
* Join that table with the `storeSize` table using `StoreId` as key. 
* Plot the the number of distinct items against the store size. 
* Add a linear regression line and smoother to the plot. 
* Can you write everything above as a single chain of commands blending the pipe operator `%>%` 
with the ggplot2 `+` concatenation of commands?

* Construct a plot of `NormalSale` against `size` for `ItemId == "7311171003510"` and
`Week == 9`. (You have to make a join!).
* Use `facet_wrap` to get all plots for this item and for the different weeks
* Use `facet_grid` to get all plots for all combinations of week and item in `fishItems`.
* Let the second axis be "free", so that it may differ for different items. Try square-root transforming the second axis as well. 

## Exercise 1D: Advanced customization {#1D}

In this exercise, you are not asked to make new chunks, but instead to modify 
the existing chunks and chunk options to customize the results.  

* Clean your R Markdown document, check the organization of chunks and give them names. 
* Change some of the plot axis labels to make them more informative. 
* Increase the width and height of the last figure above. 
* Move the color legend for one plot to the top. Then set the global theme 
option so that the legend is placed above all figures.  
* Add cache functionality to all chunks via a global option. Use auto dependencies.
* Add interactive plot / selector of items.


